---
title: 'Problem 5: Giving Data the Boot'
author: "Rakesh Vishwabrahmana"
output:
  word_document: default
  html_document: default
---



## Question 1

### Solution:
The sample mean, median, standard deviation, first quartile, geometric mean, harmonic mean, and median absolute deviation of given measurements are given in the summary.
```{r}
x <- c(207,202,169,211,191,212,108,92,186,203,126,184,206,177,164,53,190)
n <- length(x)
x_mean <- mean(x)
x_med <- median(x)
x_sd <- sd(x)
q1 <- function(x) as.vector(quantile(x, 0.25))
x_q1 <- q1(x)
gm <- function(x) exp(mean(log(x)))
x_gm <- gm(x)
hm <- function(x) 1/mean(1/x)
x_hm <- hm(x)
x_mad <- mad(x)
sum_stat <- rbind(x_mean,x_med,x_sd,x_q1,x_gm,x_hm,x_mad)
rownames(sum_stat) <- c("Mean","Median","Standard Deviation","First Quartile",
                      "Geometric Mean","Harmonic Mean","Median Absolute Deviation")
colnames(sum_stat) <- "Estimate"
sum_stat
```

## Question 2

### Solution:
The histogram of the data and a normal quantile-quantile plot is given below.
```{r}
par(mfrow=c(1,2))
hist(x)
qqnorm(x)
qqline(x)
```

## Question 3

### Solution:
In order to ascertain whether the given data show or not a serious deviation from normality, we used Shapiro-Wilk’s method.
```{r}
norm_test <- shapiro.test(x)
norm_test
```
The *p-value* of the test is `r round(norm_test$p.value,4)`. Since $p_{value}<0.05$ which implies that the data are significantly differ from normal distribution.

 * We observed from the above figure, four data points (values <150) might be an outlier and therefore resulting distribution of the given data deviates from normal.

## Question 4

### Solution:
The 95% confidence intervals on the population mean, median,standard deviation, first quartile, geometric mean, harmonic mean, and median absolute deviation using a Studentized bootstrap is given Table 2.
```{r}
boot_CI <- function(x, stat_fun, n_boot, percnt_CI) {
stat <- stat_fun(x)
n <- length(x)
boot_x <- sample(x, n*n_boot, replace = TRUE)
boot_mat <- matrix(boot_x, nrow = n, ncol = n_boot)
boot_stat <- apply(boot_mat, 2, stat_fun)
del_star <- boot_stat - stat
pct_diff <- (1 - percnt_CI) / 2
del <- as.vector(quantile(del_star, c(pct_diff, 1-pct_diff)))
ci <- stat - c(del[2], del[1])
return(ci)
}
boot_mean <- boot_CI(x, stat_fun = mean, n_boot = 1000, percnt_CI = 0.95)
boot_med <- boot_CI(x, stat_fun = median, n_boot = 1000, percnt_CI = 0.95)
boot_sd <- boot_CI(x, stat_fun = sd, n_boot = 1000, percnt_CI = 0.95) 
boot_q1 <- boot_CI(x, stat_fun = q1, n_boot = 1000, percnt_CI = 0.95) 
boot_gm <- boot_CI(x, stat_fun = gm, n_boot = 1000, percnt_CI = 0.95) 
boot_hm <- boot_CI(x, stat_fun = hm, n_boot = 1000, percnt_CI = 0.95) 
boot_mad <- boot_CI(x, stat_fun = mad, n_boot = 1000, percnt_CI = 0.95)
pop_ci <- rbind(boot_mean, boot_med, boot_sd, boot_q1, boot_gm, boot_hm, boot_mad)
colnames(pop_ci) <- c("$L_{CI}$","$U_{CI}$")
rownames(pop_ci) <- c("Mean","Median","Standard Deviation","First Quartile",
                      "Geometric Mean","Harmonic Mean","Median Absolute Deviation")
pop_ci
```
